---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

The data file is in the working directory as a compressed file. We first have to unzip the file `activity.zip`.
It generates the file `activity.csv`.

Then, we can load the data thanks to the read.csv() command.

The call to the `summary` function should show that the data are correctly loaded.

```{r echo=TRUE}
unzip("activity.zip")
data <- read.csv("activity.csv")
summary(data)
```

As the `date` field is a factor, we transform into a date thanks to the `lubridate` package.

```{r echo=TRUE}
library(lubridate)
data$date<-ymd(as.character(data$date))
```

## What is mean total number of steps taken per day?

In order to aswer this question, we first have to sum the number of steps for each day.

`temp` is a dataframe which accumulates the number of steps for each day.
```{r echo=TRUE}
temp<-aggregate(data$steps, by=list(data = data$date),  FUN = sum)
print(temp)
```

Let's plot a histogram for the number of steps in each day.
```{r echo=TRUE}
hist(temp$x, xlab="number of steps", main="Histogram of the number of steps by day")
```


Then, we have to compute the mean and the median, without taking into account NA values.
```{r echo=TRUE}
mean(temp$x, na.rm=TRUE)
median(temp$x, na.rm=TRUE)
```


## What is the average daily activity pattern?

Let us now show a time serie plot of the average number of steps for each five minute interval.

First of all, We have to aggregate the number of steps for each five minute interval.

```{r echo=TRUE}
temp<-aggregate(data$steps, by=list(interval=data$interval), FUN=mean, na.rm = TRUE)
temp
plot(temp$interval,temp$x,type="l",ylab="Average number of steps", xlab="Five minute interval of the day")
```

Let us find the interval for the maximum average number of steps.

```{r echo=TRUE}
m <- max(temp$x);
m
maxinterval<-temp[temp$x==m,]$interval
maxinterval
temp[temp$interval==maxinterval,]
```


## Imputing missing values

```{r echo=TRUE}
summary(data)
sum(is.na(data$steps))
```

Let us copy the original data then fill the missing value with the avera missing value for that interval. 


## Are there differences in activity patterns between weekdays and weekends?
